Cover letter for the original IJRR submission

### 2017-05-09 body/intro.tex
nicolas modifs

%~ Planning legged robot locomotion remains an open issue.
\IEEEPARstart{L}{egged} robots move by sequentially creating contacts with the environment.
After years of research, such robots can autonomously walk on flat ground, but struggle to navigate in more complex environments, e.g. to climb stairs using a handrail.
In such a case a humanoid robot must plan the contacts it musts create with its feet and hands. Computing such a movement is an open motion-planning problem called acyclic contact planning.
Most of the complexity of this problem lies in the choice of the sequence of contacts, i.e. the underlying decomposition of the trajectory into contact phases where specific points of the robot body are exerting forces on specific locations of the environment.
%
Once the contact plan is known, efficient approaches exist to generate a dynamically feasible motion~\citep{Carpentier2016}.

%% However in general
%% the contact plan is manually designed by a time consuming trial-and-error procedure. Automatically addressing the acyclic contact planning problem is thus a pre-requisite to the deployment of robots in cluttered, human-centered environments, which comprise constraining obstacles (stairs...).

In the specific (and simple) case of gaited biped locomotion on flat ground, choosing the effector with which to create a contact is trivial because walking follows a cyclic pattern (left foot always follows right foot).
Efficient tools such as the capture point~\citep{Pratt2006} can be used to compute the next contact location. 
%
The general case of planning contact sequences is more challenging.
At any given time a contact results of a combinatorial discrete choice for both the effector and contact surface, and a continuous choice for the contact location.
Any choice of contact then kinematically and dynamically constrains the possible motion of the robot.
Furthermore there is no analytical way to verify whether this choice brings the robot one step closer to the desired goal or to a dead end, especially in the presence of obstacles.
From this property, we say that the contact manifold is foliated~\citep{simeon-manipulation-04}. 
The foliation prevents the use of efficient sampling-based planners for two reasons.
(i) First, each submanifold of the foliation has a zero measure and cannot be directly sample.
A sample is rather obtained by sampling a ``free flying'' configuration and explicitly projecting it in contact (which is a costly numerical operation).
(ii) Second, the foliated topology turns the exploration by spreading a graph of configurations (probability roadmap, rapidely exploring random trees) into an unefficient random process, where many useless nodes are sampled on parallel submanifolds.
The total algorithm complexity of classical contact planners comes both from the number of nodes of the graph that are sampled during the exploration (ii) and to the cost of the projection when sampling a new configuration (i).
%
For this reason previous contributions having demonstrated acyclic contact locomotion on a real robot are too computationally expensive~\citep{Bretl:2006:MPM:1124573.1124585}. As a result at the DARPA Robotics Challenge, the participants stated that, except for egress, the robots did not use multi contact strategies: they relied on unsafe bipedal walking to climb stairs, instead of using the provided handrails to facilitate the motion~\citep{atkensondarpa}.

%Obtaining reasonable computation times is critical for a contact planner to be of any practical use. 
Our work aims at breaking the complexity of the acyclic contact planning problem.  
To do so we deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold (i), and the combinatorics of the contact selection problem (ii). 
To that end, we introduce a reduction property that we name the reachability condition.
This condition relaxes the strict contact condition.
It induces a low-dimensional space, called the \textit{contact reachable} space, that has a Cartesian topology, i.e. it can be efficiently sampled.
Under some hypothesis, a path in the \textit{contact reachable} space can be easily mapped to the (high-dimensionnal) contact manifold. 
This approach of the problem enables us to separate the contact planning problem in two tractable subproblems.
This decoupling presents pros and cons discussed in previous related literature, summarized in the following.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

@@

For completeness, we lastly mention a new kind of approach, recently proposed in the computer graphics field \citep{hamalainen_cpbp_2015}. The authors
used black-box physics simulators to perform Model Predictive Control for a humanoid character, and managed to obtain dynamically-consistent motions
at \gls{interactive} frame rates (a method is interactive when the computation time for planning is less than the
execution time). While this new approach provides an exciting direction of research, currently the resulting motions
 do not seem applicable to real robots.

@@

Our solution belongs to the class of decoupled approaches, because we believe it is the most promising direction~\citep{DBLP:conf/iser/EscandeKMG08} to break the complexity of acyclic contact planning. Compared to previous approaches, our solution has two main novelties:

@@

This sequential approach is the key to the efficiency of our method, though it
can result in failures. However, we demonstrate empirically its interest: the high success rate and low computation times allow us to plan (and re-plan upon failure) multi-contact sequences at \gls{interactive} rates.

>>>>

%~ Planning legged robot locomotion remains an open issue.
\IEEEPARstart{L}{egged} robots move by sequentially creating contacts with the environment.
After years of research, such robots can autonomously walk on flat ground, but struggle to navigate more complex environments.
%~ where
Deciding
where
to
create a 
contact with its feet and
possibly
its
hands
is
nontrivial,
e.g. to climb stairs using a
handrail.
%~ To climb stairs using a handrail, a humanoid robot must plan the contacts to create with its feet and hands. 
%~ Computing such motion is an open motion planning problem called acyclic contact planning.

Most of the complexity of this problem lies in the contact planning, i.e. the underlying decomposition
of the trajectory into contact phases where specific points
of
the
robot body are exerting forces on specific locations
of
the
environment. Tackling this complexity is the main
objective
of this paper.

Once the contact plan is known, efficient approaches exist to generate a dynamically feasible motion~\citep{Carpentier2016}.
 %~ However in general
%~ the contact plan is manually designed by a time consuming trial-and-error procedure. Automatically addressing the acyclic contact planning problem is thus a pre-requisite to the deployment of robots in cluttered, human-centered environments, which comprise constraining obstacles (stairs...).
In the specific (and simple) case of gaited biped locomotion on flat ground, 
%~ efficient tools such as the capture point~\citep{Pratt2006} can be used to compute the next contact location. Furthermore, 
choosing the effector with which to create a contact is trivial because walking follows a cyclic pattern (the left foot always follows the right foot).
Efficient tools such as the capture point~\citep{Pratt2006} can be used to compute the next contact location.
In the general case planning complex contact interactions is extremely challenging:
At any given time a contact must be chosen between infinitely many possibilities (often a combinatorial discrete choice for the effector and contact surface, and a continuous choice for the contact location). Furthermore, a contact choice constrains kinematically and dynamically the possible motions, and there is no analytical way to verify whether this choice brings the robot one step closer to the desired goal or to a dead end, especially in the presence of obstacles; we say that the contact manifold is foliated~\citep{simeon-manipulation-04}. 
%~ Lastly, the contact manifold has a null measure and is thus impossible to sample. This restriction prevents the use of efficient sampling-based planners.
The
foliation prevents the use of efficient sampling-based
planners
for two reasons. (i) First, each submanifold of the         
foliation
has
a zero measure and cannot be directly sampled.         
A
sample
is
rather obtained by sampling a “free flying”         
configuration
and
explicitly projecting it in contact (which is a         
costly
numerical
operation).
(ii) Second, the foliated topology       
turns
the
exploration
by
spreading
a graph of configurations        
(probabilistic
roadmap,
rapidely
exploring
random trees) into      
an
unefficient
random
process,
where
many
useless nodes       
are
sampled
on
parallel
submanifolds.
The
total
algorithmic       
complexity
of
classical
contact
planners
comes
from
both the        
number
of
graph
nodes
sampled
during
exploration
(ii) and
the cost of the projection when sampling new configurations
(i).

For this reason previous contributions having demonstrated acyclic contact locomotion on a real robot are too computationally expensive~\citep{Bretl:2006:MPM:1124573.1124585}. As a result at the DARPA Robotics Challenge, the participants stated that except for egress, the robots did not use multi contact strategies: they relied on unsafe bipedal walking to climb stairs, instead of using the provided handrails to facilitate the motion~\citep{atkensondarpa}.

%~ Obtaining reasonable computation times is critical for a contact planner to be of any practical use.
 Our work aims at breaking the complexity of the acyclic contact planning problem.  To do so we deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold, and the combinatorics of the contact selection problem. First we introduce a low-dimensional space, called the \textit{contact reachable} space, that can be sampled and mapped efficiently to the contact manifold. Then, given a path computed in the \textit{contact reachable} space, we propose a deterministic algorithm to generate a contact sequence along the path.
This decoupling presents pros and cons discussed in previous related literature, summarized in the following.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

@@

%~ For completeness, we lastly mention a new kind of approach, recently proposed in the computer graphics field \citep{hamalainen_cpbp_2015}. The authors
%~ used black-box physics simulators to perform Model Predictive Control for a humanoid character, and managed to obtain dynamically-consistent motions
%~ at \gls{interactive} frame rates (a method is interactive when the computation time for planning is less than the
%~ execution time). While this new approach provides an exciting direction of research, currently the resulting motions
 %~ do not seem applicable to real robots.

@@

Our solution belongs to the class of decoupled approaches, 
i.e. 
we
propose
specific
algorithms
to
efficiently
solve
both
\Pa
and \Pb
while 
relying 
on 
state-of-the-art 
solution 
to 
$\mathcal{P}_3$
to 
obtain
the
whole
movement.
Our
main
contribution
is
the
definition        
of
a reduction property, the reachability condition.

%~ because we believe it is the most promising direction~\citep{DBLP:conf/iser/EscandeKMG08} to break the complexity of acyclic contact planning. 
Compared to previous approaches, our solution has two main novelties:

@@

The reachability condition is the key to the strict separation between $\mathcal{P}_1$ and $\mathcal{P}_2$, hence to the low complexity of our planner, though it
can result in failures. However, we demonstrate empirically its interest: the high success rate and low computation times allow us to plan (and re-plan upon failure) multi-contact sequences at \gls{interactive} rates.

### 2017-05-09 body/results.tex
nicolas modifs

\subsection{Validation of the contact plans, and comparison with previous work} \label{sec:compa}
To validate our plans, we either use the framework of \cite{Carpentier2016}, or our own implementation of a $\mathcal{P}_3$ solver (Appendix~\ref{app:optim}).
The companion video shows the obtained motions.

Few contributions provided computation times on the complete problem. %, necessary to provide comparison between our approaches. 
Furthermore, $\mathcal{P}_3$ remains an open problem in the presence of obstacles. The only valid scenarios addressed completely in previous works are thus the stair-climbing scenarios of different heights proposed by Hauser in \cite{Hauser06usingmotion}, and the table-egress scenario by Escande et al. in~\cite{DBLP:conf/iser/EscandeKMG08}, which we consider to be of similar complexity with respect to the car-egress scenario (we did not consider the stairs in the scene). Both scenarios are tested with HRP-2.

@@

To provide a completely fair comparison, it would be required to compare the methods on the same computers and exact same scenarios, which is not possible
because the source code of the other methods is often not available. Despite this observation and the fact that some results are old, we believe that the results show that our method
performs better. Indeed, a difference in technology does not justify a difference as important as the one demonstrated in Table~\ref{tab:sucess_planning}.

>>>>

\subsection{Validation of the contact plans} \label{sec:valid}
To validate our plans, we either use the framework of \cite{Carpentier2016}, or our own implementation of a $\mathcal{P}_3$ solver (Appendix~\ref{app:optim}).
The companion video shows the obtained motions.

\subsection{Comparison with previous work} \label{sec:compa}
%~ Few contributions provided computation times on the complete problem. To provide a completely fair comparison, it would be required to compare the methods on the same computers and exact same scenarios, which is not possible because 
%~ the source code of the other methods is often not available. %, necessary to provide comparison between our approaches. 
We did our best to provide a fair comparison of the computation complexity of our method with the state of the art. 
%Few contributions provided computation times on the complete problem. 
However existing benchmarks for motion planning algorithms~\cite{moll2014extensible} do not yet encompass contact planning.
Moreover, the source code of the previous methods of the state of the art is often not available.
Providing a fair comparison with the algorithms performing on the same computer and on the same scenarios is yet out of reach.
A step in this direction is the open-source release of our source code (see Appendix~\ref{app:hpp}) that allows any reader to reproduce our results.
Furthermore, $\mathcal{P}_3$ remains challenging in the presence of obstacles. The only valid scenarios addressed completely in previous works are thus the stair-climbing scenarios of different heights proposed by Hauser in \cite{Hauser06usingmotion}, and the table-egress scenario by Escande et al. in~\cite{DBLP:conf/iser/EscandeKMG08}, which we consider to be of similar complexity with respect to the car-egress scenario (we did not consider the stairs in the scene). Both scenarios are tested with HRP-2.

@@

 %~ Despite this observation and the fact that some results are old, we believe that the results show that our method
%~ performs better. Indeed, a difference in technology does not justify a difference as important as the one demonstrated in Table~\ref{tab:sucess_planning}.
